Open-Set Domain Adaptation through Self-Supervision

### 2022-02-13 paper/main.tex
minor fixes

In contrast, the presented method uses a Self-Supervised approach, already presented in 
the recent years by many other works \cite{sun2019unsupervised,saito2020universal,yue2021prototypical}.
using pretext tasks (e.g., image rotation, jigsaw puzzle \cite{CarlucciJigsaw}, mutual information (MI) \cite{selfsupfeng}, instance discrimination \cite{chen2020simple}) to learn high-level feature representation in 
source and target domains by jointly training a shared feature extractor.

@@

In the recent years different works to a Self-Supervised solution, used also in the 
presented method, engaging in learning invariant representations to the 
transformations of the input data, improving separation of classes from each other 
and from open-set samples \cite{dhamija2021selfsupervised,jia2021selfsupervised,GeoTran}.

\textbf{Self-supervised Learning} is an emerging field focused on using 
data itself as supervision for auxiliary (also called “pretext”) 
tasks that learn deep feature representations, which will hopefully be informative 
for downstream “real” tasks.
Many such auxiliary tasks have been proposed in the literature, including
image colorization \cite{zhang2016colorful},
image jigsaw puzzle [\cite{noroozi2017unsupervised}, geometric transformations \cite{dosovitskiy2015discriminative}.

Another geometric transformation task is rotation recognition \cite{OldROS}, where input images
are rotated by multiples of 90° and the network is trained to predict the rotation
angle pf each image. This pretext task has been successfully used for both anomaly
detection \cite{GeoTran} and closed-set domain adaptation \cite{SelfSupervisedXu}.

\textbf{Open Set Domain Adaptation} (OSDA) is a challenging domain adaptation 
setting which allows the existence of unknown classes on the target domain.
The term "OSDA" was first introduced by Busto and Gall \cite{firstosda}, but the 
currently accepted definition was introduced by Saito \etal \cite{saito2018open} considering the target as 
containing all the source categories and additional set of private categories 
that should be considered unknown. They have presented Open Set
Back-Propagation (OSBP) \cite{saito2018open} as an adversarial method 
to increase the prediction variances so that the generator can choose if accept Ornot a target saple.
Separate To Adapt (STA) \cite{sat} train a multi-binary classifier to progressively 
separate the samples of unknown and known classes.
Universal Adaptation Network (UAN) \cite{uan}, originally proposed for the universal 
domain adaptation setting that is a superset of OSDA, quantify the sample-level 
transferability and recognize the unknown samples based on it.
Attract or Distract (AoD) \cite{feng2019attract} refine the decision by using 
metric learning to reduce the intra-class distance in known classes and push 
the unknown class away from the known classes.
Jing \etal \cite{Jing_Li_Zhu_Ding_Lu_Yang_2021} have recently proposed a method
to recognize the unknown samples according to the centroid deviation angles and
employing the statistical Extreme Value Theory to recognize the unknown 
samples that are misclassified into known classes.

@@

{\it Multi-head rotation classifier:}
alternatively, 
instead of having a single rotation head predicting the rotation of a sample regardless of its semantic class, 
we also try using a different head for each known class $\in \mathcal{C}_s$, each one responsible of predicting the rotation
of the images belonging to that semantic class.
This variation can mitigate the problem of trying to predict the rotation of a larger number
of semantic classes. Infact, as the number of semantic classes grows,
the problem of predicting the relative orientation becomes more difficult.
The application of the rotation recognition pre-text task allows to effectively favor and force the model to learn
domain-independent patterns, which are crucial to perform the novelty detection in a cross-domain fashion and, moreover,
to successfully perform the domain alignment.
To provide an explanation of why this applies, we can think that a rotation classifier needs to focus on discriminative patterns to
successfully perform the rotation predictions, such as shapes, edges, and high-level object relative position like the position of the eyes
w.r.t nose.
The method and its effectiveness is further illustrated in \cite{OldROS},
and further improvements will be discussed in \ref{sec:future_work}.

>>>>

In contrast, the presented method uses a Self-Supervised approach, already presented in 
the recent years by many other works \cite{sun2019unsupervised,saito2020universal,yue2021prototypical}
using pretext tasks (e.g., image rotation, jigsaw puzzle \cite{CarlucciJigsaw}, mutual information (MI) \cite{selfsupfeng}, instance discrimination \cite{chen2020simple}) to learn high-level feature representation in 
source and target domains by jointly training a shared feature extractor.

@@

In the recent years different works have landed to a Self-Supervised solution, used also in the 
presented method, engaging in learning invariant representations to the 
transformations of the input data, improving separation of classes from each other 
and from open-set samples \cite{dhamija2021selfsupervised,jia2021selfsupervised,GeoTran}.

\textbf{Self-supervised Learning} is an emerging field focused on using 
data itself as supervision for auxiliary (also called “pretext”) 
tasks that learn deep feature representations, which will hopefully be informative 
for downstream “real” tasks.
Many such auxiliary tasks have been proposed in the literature, including
image colorization \cite{zhang2016colorful},
image jigsaw puzzle \cite{noroozi2017unsupervised}, geometric transformations \cite{dosovitskiy2015discriminative}.

Another geometric transformation task is rotation recognition \cite{OldROS}, where input images
are rotated by multiples of 90° and the network is trained to predict the rotation
angle of each image. This pretext task has been successfully used for both anomaly
detection \cite{GeoTran} and closed-set domain adaptation \cite{SelfSupervisedXu}.

\textbf{Open Set Domain Adaptation} (OSDA) is a challenging domain adaptation 
setting which allows the existence of unknown classes on the target domain.
The term "OSDA" was first introduced by Busto and Gall \cite{firstosda}, but the 
currently accepted definition was introduced by Saito \etal \cite{saito2018open} considering the target as 
containing all the source categories and additional set of private categories 
that should be considered unknown. They have presented Open Set
Back-Propagation (OSBP) \cite{saito2018open} as an adversarial method 
to increase the prediction variances so that the generator can choose if accept or not a target saple.
Separate To Adapt (STA) \cite{sat} train a multi-binary classifier to progressively 
separate the samples of unknown and known classes.
Universal Adaptation Network (UAN) \cite{uan}, originally proposed for the universal 
domain adaptation setting that is a superset of OSDA, quantify the sample-level 
transferability and recognize the unknown samples based on it.
Attract or Distract (AoD) \cite{feng2019attract} refine the decision by using 
metric learning to reduce the intra-class distance in known classes and push 
the unknown class away from the known classes.
Jing \etal \cite{Jing_Li_Zhu_Ding_Lu_Yang_2021} have recently proposed a method
to recognize the unknown samples according to the centroid deviation angles and
employing the statistical Extreme Value Theory to recognize the unknown 
samples that are misclassified into known classes.

@@

{\it Multi-head rotation classifier:}
alternatively, 
instead of having a single rotation head predicting the rotation of a sample regardless of its semantic class, 
we also try using a different head for each known class $\in \mathcal{C}_s$, each one responsible of predicting the rotation
of the images belonging to that semantic class.
This variation can mitigate the problem of trying to predict the rotation of a larger number
of semantic classes. Infact, as the number of semantic classes grows,
the problem of predicting the relative orientation becomes more difficult.

The application of the rotation recognition pre-text task allows to effectively favor and force the model to learn
domain-independent patterns, which are crucial to perform the novelty detection in a cross-domain fashion and, moreover,
to successfully perform the domain alignment.
To provide an explanation of why this applies, we can think that a rotation classifier needs to focus on discriminative patterns to
successfully perform the rotation predictions, such as shapes, edges, and high-level object relative position like the position of the eyes
w.r.t nose.
The method and its effectiveness is further illustrated in \cite{OldROS},
and further improvements will be discussed in \ref{sec:future_work}.
